Efficiently Linking Documents to Alignment Databases -- A Prototype-Based Exploration

### 2025-09-26 doc/aligments2025/doc.tex
more abstract writing

\documentclass{ceurart}
\usepackage[show]{ed}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
}

@@

The sTeX ecosystem, which uses diverse semantic markup to drive
learning support systems,
faced a similar challenge, which was recently addressed by the snify system.
In this paper, we present a prototypical extension of the snify system
for efficient alignment that is independent of the sTeX ecosystem.
Concretely, we have added the option to use WikiData as the concept database,
rudimentary support for annotating formulae,
the ability to annotate HTML documents,
and a browser-based user interface, which
helps with the annotation of HTML documents.

@@

As WikiData also has formula notations
for many concepts, including \LaTeX\ notations like \lstinline.\mathbb{N}.,
we can also use the same workflow to align formula fragments.
Such formula alignment is rather limited:
operators can be aligned, but their arguments are not marked, and
some operators, like the invisible multiplication in ``$ax$'',
are not represented by a glyph at all.
In sTeX, the solution is essentially to write down
a semantic representation of the formula and generate the presentation from it.
A grammar-based approach to generate the semantic representation
for ``conventional'' \LaTeX\ formulae
in collaboration with a human annotator was presented in\ednote{TODO: Luca's work}.
For alignments, this approach could be considered too invasive.

@@

>>>>

\documentclass{ceurart}
\usepackage[show]{ed}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily,
}

@@

The sTeX ecosystem, which uses diverse semantic markup to drive
learning support systems,
faced a similar challenge, which was recently addressed by the snify system\cite{KohSch:safc25}.
In this paper, we present a prototypical extension of the snify system
for efficient alignment that is independent of the sTeX ecosystem.
Concretely, we have added the option to use WikiData as the concept database,
rudimentary support for annotating formulae,
the ability to annotate HTML documents,
and a browser-based user interface, which
helps with the annotation of HTML documents.

@@

As WikiData also has formula notations
for many concepts, including \LaTeX\ notations like \lstinline.\mathbb{N}.,
we can also use the same workflow to align formula fragments.
Such formula alignment is rather limited:
operators can be aligned, but their arguments are not marked, and
some operators, like the invisible multiplication in ``$ax$'',
are not represented by a glyph at all.
In sTeX, the solution is essentially to write down
a semantic representation of the formula and generate the presentation from it.
A grammar-based approach to generate the semantic representation
for ``conventional'' \LaTeX\ formulae
in collaboration with a human annotator was presented in\cite{VreWelKam:tsmmdui24}.
For alignments, this approach could be considered too invasive.

@@

% \printbibliography

\bibliography{references}
